---
title: "Exploratory Graphs and Maps"
author: "McEwen Khundi"
date: "24 March 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r packages}
library(here)
library(tidyverse)
library(sf)
library(ggspatial) 
library(scales)
library(viridis)
library(arsenal)
library(ggthemes)
#devtools::install_github("petermacp/BlantyreTBEpi")
library(BlantyreTBEpi)
```

Read data 
```{r}
scaleFUN <- function(x) sprintf("%.2f", x)

malawi = readRDS(file = here("data/malawi.rds"))
malawi_lakes = readRDS(file = here("data/malawi_lakes.rds"))
africa = readRDS(file = here("data/africa.rds"))

data("hsas", package="BlantyreTBEpi")
data("clinics", package="BlantyreTBEpi")
data("hsas", package="BlantyreTBEpi")
data("dat", package="BlantyreTBEpi")
data("x01_epal2", package="BlantyreTBEpi")
data("cnrs", package="BlantyreTBEpi")
data("cnrs", package="BlantyreTBEpi")
```

Map theme
Create a base theme function for the maps.
```{r}
map_light <- function(...) {
  theme_minimal() +
  theme(
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_line(color = "#ebebe5", size = 0.2),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", color = NA), 
    panel.background = element_rect(fill = "white", color = NA), 
    legend.background = element_rect(fill = "white", color = NA),
    panel.border = element_blank(),
    ...
  )
}

```

## 1. Map of countries in Africa
```{r}
fg1a <- ggplot() +
  geom_sf(data=africa, aes(fill = bt_label)) +
  scale_fill_manual(values = alpha(c("#DDDDDD", "#FF851B")),guide=FALSE) +
  labs(y = "Latitude", 
       x = "Longitude",
       title = "Map of countries in Africa",
       subtitle = "Malawi is marked orange") +
  theme_light() +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "tr", style = north_arrow_fancy_orienteering)

 fg1a
 ggsave("africa_countries.png", height = 10, width = 8, dpi=300, 
       path = here("figures"))
```


## 1. Map of administrative districts of Malawi
```{r}
fg1 <- ggplot() +
  geom_sf(data=malawi, aes(fill = bt_label)) +
  scale_fill_manual(values = alpha(c("#DDDDDD", "#FF851B")),guide=FALSE) +
  labs(y = "Latitude", 
       x = "Longitude",
       title = "Map of administrative districts of Malawi",
       subtitle = "Blantyre is marked orange") +
  geom_sf(data=malawi_lakes[malawi_lakes$NAME=="LAKE NYASA",], fill = "light blue", color = "blue") +
  theme_light() +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "tr", style = north_arrow_fancy_orienteering)

 fg1
 ggsave("malawi_districts.png", height = 10, width = 8, dpi=300, 
       path = here("figures"))
```

## 2. Map of 315 community catchment areas in Blantyre
```{r}

fg2 <- ggplot() +
  geom_sf(data=hsas) +
  geom_sf(data=clinics, shape = 17) +
  labs(y = "Latitude", 
       x = "Longitude",
       title = "Map of 315 community catchment areas in Blantyre",
       subtitle = "Each cluster has a  responsible CHW\n The triangles represent TB registration clinics") +
  theme_light() +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "tr", style = north_arrow_fancy_orienteering)
 
fg2 
 ggsave("hsa_areas_clinics.png", height = 10, width = 8, dpi=300, 
       path = here("figures")) 
```

## 3. Map of the population of Blantyre
```{r}

 fg3 <- ggplot() +
  geom_sf(data=dat, aes(fill = total) ) +
  scale_fill_viridis(option = "magma", direction = -1,name ="Count", labels=comma) +
  geom_sf(data=clinics, shape = 17) +
  labs(y = "Latitude", 
       x = "Longitude",
       title = "Total population of Blantyre") +
  theme_light() +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "tr", style = north_arrow_fancy_orienteering)

 fg3  
 ggsave("total_pop.png", height = 10, width = 8, dpi=300, 
       path = here("figures")) 
```

## 4. Map of the population density of Urban Blantyre
```{r}
 fg4 <- ggplot() +
  geom_sf(data=dat, aes(fill = popdens) ) +
  scale_fill_viridis(option = "magma", direction = -1,
                     name = expression ("Population\nper\nsquare km"), labels=comma) +
  geom_sf(data=clinics, shape = 17) +
  labs(y = "Latitude", 
       x = "Longitude",
       title = "Population density of Blantyre") +
  theme_light() +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "tr", style = north_arrow_fancy_orienteering)
 
fg4 
 ggsave("pop_density.png", height = 10, width = 8, dpi=300, 
       path = here("figures")) 
```

## 5 Mean number of people per household in Blantyre
```{r}
fg5 <-  ggplot() +
  geom_sf(data=dat, aes(fill = mean_people_per_hh)) +
    scale_fill_viridis(option = "magma", direction = -1,
                     name = expression ("Mean number\nof people\nper household"), labels=scaleFUN) +
  geom_sf(data=clinics, shape = 17) +
  labs(y = "Latitude", 
       x = "Longitude",
       title = "Mean number people per household in Blantyre") +
  theme_light() +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "tr", style = north_arrow_fancy_orienteering)
 
fg5 
 ggsave("mean_household.png", height = 10, width = 8, dpi=300, 
       path = here("figures")) 
```

## 6. Poverty distribution
Plot the estimaed mean proportion of the population living on less than $2 per day. [Note this is estimated from Worldpop data]

```{r}
fg6 <-  ggplot() +
  geom_sf(data=dat, aes(fill = mwi11povcons200)) +
     scale_fill_viridis(option = "magma", direction = -1,
                     name = expression ("Mean percentage\nof people\nliving on\nless than\nUS $2 per day"),
                     labels=comma, limits = c(0,100)) +
  geom_sf(data=clinics, shape = 17) +
  labs(y = "Latitude", 
       x = "Longitude",
       title = "Mean percentage of people living on less than\n US $2 per day in Blantyre") +
  theme_light() +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "tr", style = north_arrow_fancy_orienteering)
 
fg6 
 ggsave("mean_poverty.png", height = 10, width = 8, dpi=300, 
       path = here("figures")) 
```

## 7. Distance to nearest TB registration site

```{r}
fg7 <-  ggplot() +
  geom_sf(data=dat, aes(fill = clinic_distance)) +
     scale_fill_viridis(option = "magma", direction = -1,
                     name = expression ("Distance (m)\nto nearest\nTB registration\ncentre"),
                     labels=comma,
                     limits = c(0,10000)) +
  geom_sf(data=clinics, shape = 17) +
  labs(y = "Latitude", 
       x = "Longitude",
       title = "Distance (m) to nearest TB registration centre in Blantyre") +
  theme_light() +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "tr", style = north_arrow_fancy_orienteering)
 
fg7 
 ggsave("clinic_distance.png", height = 10, width = 8, dpi=300, 
       path = here("figures")) 
```

## 8. Male to female adult ratio
```{r}
fg8 <-  ggplot() +
  geom_sf(data=dat, aes(fill = m_f_adult_ratio)) +
     scale_fill_viridis(option = "magma", direction = -1,
                        name = expression ("Adult\nMale:Female\nratio"),
                     labels=scaleFUN) +
  geom_sf(data=clinics, shape = 17) +
  labs(y = "Latitude", 
       x = "Longitude",
       title = "Adult male to female ratio in Blantyre") +
  theme_light() +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "tr", style = north_arrow_fancy_orienteering)

fg8  
 ggsave("male_female.png", height = 10, width = 8, dpi=300, 
       path = here("figures")) 
```

## 9. Percentage of HSA cluster population who are adult
```{r}
fg9 <- ggplot() +
  geom_sf(data=dat, aes(fill = prop_adults)) +
     scale_fill_viridis(option = "magma", direction = -1,
                     name = expression ("Percent\naged\n15 years\nor older"),
                     labels=comma) +
  geom_sf(data=clinics, shape = 17) +
  labs(y = "Latitude", 
       x = "Longitude",
       title = "Percent aged 15 years or older in Blantyre") +
  theme_light() +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "tr", style = north_arrow_fancy_orienteering)
  
fg9
 ggsave("prop_adults.png", height = 10, width = 8, dpi=300, 
       path = here("figures")) 
```

## 10. Sputum smear positive to negative ratio
```{r}
#Add in a term for ratio of smearpositive to smearnegative cases

dat <- dat %>%
  mutate(smr_ratio = n_smearpos_cases/(n_cases-n_smearpos_cases))

dat$smr_ratio[is.nan(dat$smr_ratio)] <- 1
dat$smr_ratio[is.infinite(dat$smr_ratio)] <- 1


fg10 <- ggplot() +
  geom_sf(data=dat, aes(fill = smr_ratio)) +
     scale_fill_viridis(option = "magma", direction = -1,
                     name = expression ("Smr +ve:-ve\nratio"),
                     labels=scaleFUN)  +
  geom_sf(data=clinics, shape = 17) +
  labs(y = "Latitude", 
       x = "Longitude",
       title = "Sputum smear positive to negative ratio in Blantyre") +
  theme_light() +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "tr", style = north_arrow_fancy_orienteering)

fg10
 ggsave("spear_posneg.png", height = 10, width = 8, dpi=300, 
       path = here("figures")) 
```

## 11. Table 1: Characteristics of TB cases
Compare characteristics of cases located within and outside of HSA catchment areas.
```{r,  results="asis"}
x01_epal2$year <- fct_drop(x01_epal2$year)


table1 <- tableby(resident ~ year + sex + age + tbclass + smearstatus + hivstatus + lab_smear + lab_culture + lab_id + any_micro_confirm, data = x01_epal2)

summary(table1)
```

## 12. TB case notification rates
Plot overall TB case notification rates.

```{r}
cnr1 <- ggplot() +
  geom_sf(data=dat) +
  geom_sf(data=cnrs, aes(fill=cnr)) + 
  geom_sf(data=clinics, shape=17, colour="#22211d") +
   map_light() +
  scale_fill_distiller(palette = "Spectral",
                     name = "",
                     labels=comma,
                     limits = c(0,999)) +
  labs(x = NULL, 
       y = NULL,
       title = "TB case notification rates") +
  facet_grid(~year, drop = FALSE) +
  theme(panel.background = element_rect(fill = "white", color = "#22211d" )) 

cnr1
ggsave("case_notifications.png", width=12, height=6, dpi=300, 
       path = here("figures")) 
```

Now plot microbiologically-confirmed TB case notification rate
```{r}
cnr2 <- ggplot() +
  geom_sf(data=dat) +
  geom_sf(data=cnrs, aes(fill=micro_cnr)) + 
  geom_sf(data=clinics, shape=17, colour="#22211d") +
  map_light() +
  scale_fill_distiller(palette = "Spectral",
                     name = "",
                     labels=comma,
                     limits = c(0,500)) +
  labs(x = NULL, 
       y = NULL) +
  facet_grid(~year, drop = FALSE) +
  theme(panel.background = element_rect(fill = "white", color = "#22211d" ))

cnr2

ggsave(file="micro_cnrs.png", width=12, height=6, dpi=300, 
       path = here("figures"))
```







